\papertitle

### 2013-02-09 manuscript.tex
moving headings
shifted some text

The literature on cooperative control of MAVs is similarly limited to larger
vehicles or analytical explorations. Hyams et al., Jung et al. and Nebot et
al. all demonstrated cooperative visual servoing for heterogeneous
UGVs~\cite{Hyams1999Cooperative}\cite{Jung1998Range}\cite{Nebot2003Agents}.
Luo, et al. and Mehta, et al. analyze UAV-UGV cooperation from a modeling
perspective~\cite{Luo2011Air}\cite{Mehta2006Adaptive}. Rudol et al.
demonstrated a similar cooperative visual servoing system with UGVs and
MAVs~\cite{Rudol2008Micro}, but the vehicles used were orders of magnitude
larger than the lightweight systems we target, and so were able to take
advantage of high-performance processors and specialized cameras. Stirling et
al. describe a method for cooperation amongst quadrotor MAVs to navigate an indoor
environment~\cite{SwarmStirling}. Stationary MAVs provide external sensing for
navigation through the environment, and the network can be extended by using
the exploring MAVs for sensing in undiscovered areas. Their implementation
relies solely on local sensing, although our size range requires that the
actual environment sensing be external to the MAV. Additionally, they do not
consider time constraints in their planning, whereas our MAV has a minimum
forward speed and turning radius.

The small size of lightweight air and ground vehicles necessitates 
specialization, but it also pays dividends in agility. A group of specialized 
heterogeneous, lightweight robots can navigate more challenging terrains and 
interact with their environments in ways that larger, monolithic vehicles 
cannot. Despite this advantage, in order to collectively equal the sensing and 
computational capabilities of larger robots, small robots will have to interact 
so that multiple robots can benefit from the specialized capabilities of 
cooperating vehicles. In this paper we consider both experimentally, and in 
simulation, the performance of the externally-directed ornithopter to 
demonstrate how millirobotic system can benefit from cooperation.

@@

The high pitch angles at which flapping-wing MAVs regularly operate 
distinguish them from other MAVs. Baek demonstrated control on a similar 
vehicle using PID and an Euler angle parameterization of orientation~\cite{baek:tracking}. 
However, Euler angle parameterizations of vehicle orientation suffer from 
singularities around high pitch angles. So we instead represent the vehicle 
orientation with quaternions. The quaternion representation allows us to 
represent relative body and world coordinate angular displacements compactly 
with quaternion multiplication~\cite{hamilton1866elements}.

Our implementation of quaternion-based control is similar to 
Knoebe's~\cite{knoebe:quatcontrol}:
Given a reference pose represented in quaternion form $q_r$, we define the 
error as the rotation $q_e$ required to reach the reference pose from the 
estimated pose $q$:
\begin{align}
\label{eq:quat_error}
q_e &= q^{\prime}\cdot q_r \\
		&= q_{e,w} + q_{e,x}\hat{i} + q_{e,y}\hat{j} + q_{e,z}\hat{k} \notag 
\end{align}
We then convert this error quaternion to an angle axis representation and 
use it as input to the PID controller (Equations~\ref{eq:quat_linearize} and~\ref{eq:quat_angle}).
We perform these calculations on-board, with trigonometric lookup tables for 
real-time performance.
\begin{align}
\label{eq:quat_linearize}
Q_e &= \alpha\cdot q_e/\sin(\alpha /2) \\
		&= \alpha_{x}\hat{i} + \alpha_{y}\hat{j} + \alpha_{z}\hat{k} \notag
\end{align}
\begin{equation}
\label{eq:quat_angle}
\alpha = 2\arccos(q_{e,w})
\end{equation}

@@

%----------------------------------------------------------------------------%
\vspace{0.28in}\subsection{Ornithopter Attitude Control}
\label{sec:model_attitude}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}
We have demonstrated cooperative guidance of a 13 gram flapping-wing MAV with
a lightweight base station. In addition, we developed a model of
the cooperative system using system identification methods on the individual
system capabilities, and composing it with known system interactions. The
corroboration between our simulated and experimental results shows that our
simplified modeling approach is useful for predicting the performance of
cooperative systems at low computational cost. Our model could also be used to
determine when to begin control with high probability of success, even though
we lack information about the complete pose of the MAV.

>>>>

The literature on cooperative control of MAVs is similarly limited to larger
vehicles or analytical explorations. Hyams et al., Jung et al. and Nebot et
al. all demonstrated cooperative visual servoing for heterogeneous
UGVs~\cite{Hyams1999Cooperative}\cite{Jung1998Range}\cite{Nebot2003Agents}.
Luo, et al. and Mehta, et al. analyze UAV-UGV cooperation from a modeling
perspective~\cite{Luo2011Air}\cite{Mehta2006Adaptive}. Rudol et al.
demonstrated a similar cooperative visual servoing system with UGVs and
MAVs~\cite{Rudol2008Micro}, but the vehicles used were orders of magnitude
larger than the lightweight systems we target, and so were able to take
advantage of high-performance processors and specialized cameras. Stirling et
al. describe a method for cooperation amongst quadrotor MAVs to navigate an indoor
environment~\cite{SwarmStirling}. Stationary MAVs provide external sensing for
navigation through the environment, and the network can be extended by using
the exploring MAVs for sensing in undiscovered areas. Their implementation
relies solely on local sensing, although our size range requires that the
actual environment sensing be external to the MAV. Additionally, they do not
consider time constraints in their planning, whereas our MAV has a minimum
forward speed and turning radius.
\looseness -1
The small size of lightweight air and ground vehicles necessitates 
specialization, but it also pays dividends in agility. A group of specialized 
heterogeneous, lightweight robots can navigate more challenging terrains and 
interact with their environments in ways that larger, monolithic vehicles 
cannot. Despite this advantage, in order to collectively equal the sensing and 
computational capabilities of larger robots, small robots will have to interact 
so that multiple robots can benefit from the specialized capabilities of 
cooperating vehicles. In this paper we consider both experimentally, and in 
simulation, the performance of the externally-directed ornithopter to 
demonstrate how millirobotic system can benefit from cooperation.

@@

The high pitch angles at which flapping-wing MAVs regularly operate 
distinguish them from other MAVs. Baek demonstrated control on a similar 
vehicle using PID and an Euler angle parameterization of orientation~\cite{baek:tracking}. 
However, Euler angle parameterizations of vehicle orientation suffer from 
singularities around high pitch angles. So we instead represent the vehicle 
orientation with quaternions. The quaternion representation allows us to 
represent relative body and world coordinate angular displacements compactly 
with quaternion multiplication~\cite{hamilton1866elements}.
\looseness -1
Our implementation of quaternion-based control is similar to 
Knoebe's~\cite{knoebe:quatcontrol}:
Given a reference pose represented in quaternion form $q_r$, we define the 
error as the rotation $q_e$ required to reach the reference pose from the 
estimated pose $q$:
\begin{align}
\label{eq:quat_error}
q_e &= q^{\prime}\cdot q_r \\
		&= q_{e,w} + q_{e,x}\hat{i} + q_{e,y}\hat{j} + q_{e,z}\hat{k} \notag 
\end{align}
We then convert this error quaternion to an angle axis representation and 
use it as input to the PID controller (Equations~\ref{eq:quat_linearize} and~\ref{eq:quat_angle}).
We perform these calculations on-board, with trigonometric lookup tables for 
real-time performance.
\begin{align}
\label{eq:quat_linearize}
Q_e &= \alpha\cdot q_e/\sin(\alpha /2) \\
		&= \alpha_{x}\hat{i} + \alpha_{y}\hat{j} + \alpha_{z}\hat{k} \notag
\end{align}
\begin{equation}
\label{eq:quat_angle}
\alpha = 2\arccos(q_{e,w})
\end{equation}

@@

%----------------------------------------------------------------------------%
\subsection{Ornithopter Attitude Control}
\label{sec:model_attitude}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions and Future Work}
We have demonstrated cooperative guidance of a 13 gram flapping-wing MAV with
a lightweight base station. In addition, we developed a model of
the cooperative system using system identification methods on the individual
system capabilities, and composing it with known system interactions.

\noindent The corroboration between our simulated and experimental results shows that our
simplified modeling approach is useful for predicting the performance of
cooperative systems at low computational cost. Our model could also be used to
determine when to begin control with high probability of success, even though
we lack information about the complete pose of the MAV.
